---
title: "Reynold's Creek Experimental Watershed: Single Point Simulations"
author: "Katie Murenbeeld"
date: "6/27/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

Reynold's Creek Experimental Watershed (RCEW)...(a little about the location and CZO).

About the Community Earth Systems Model (CESM), Community Terrestrial Systems Model (CTSM) and Community Land Model (CLM) explain how different although terms are used almost interchangably. Model simulations. Include a link to the escomp CLM5.0 and the CLM4.5 User's Guides. 

The purpose of this technical note is to familiarize a LEAF lab user with setting up and running a single point simulation at RCEW with meterological data from one of the AmeriFlux towers on the site. In this document, I am assuming that you have a small or large allocation on the National Center for Atmospheric Research's (NCAR) supercomputer Cheyenne. At this time CLM is not ported on the high performance computer (HPC) R2 at Boise State, but this is a goal to complete soon. 

```{r clm_fig01, echo=FALSE, eval=FALSE}

```

## File Structure and Cloning CTSM

### File Structure

If you don't have a working directory set up already on Cheyenne, do so now. It is a best practice to 

```{bash set_dir, echo=TRUE, eval=FALSE}
cd /glade/work/$USER/ # go to your work directory on Glade
mkdir git             # create a new folder that will be the parent for future projects using CLM/CTSM
cd git                # go to the newly created git folder
```

### Cloning CTSM

From your newly created git folder, clone the CLM/CTSM code. The following code will point to the latest CTSM release:

```{bash git_clone, echo=TRUE, eval=FALSE}
# Check out the latest release branch and create a new branch for your project

git clone -b release-clm5.0 https:github.com/ESCOMP/CTSM.git ctsm_rcew 
cd ctsm_rcew   # A new directory was created for your project
git branch   # Check to see the branches you have now created. There should be a master (or main) as well as a ctsm_rcew branch. The ctsm_rcew branch should have a star next to it.
ls   # Explore the contents of the new directory
```

Notice that there is a folder within your project direcotry named manage_externals. This contains a very important script, checkout_externals, which downloads all of the components required for building and running CTSM. These include: CLM (land model component), CISM (sea-ice component), RTM (river routing component), MOSART (river routing component), CIME (contains scripts and tools for running CESM), CMEPS (Earth prediction option of CIME), FATES (dynamic vegetation option of CLM), and PTCLM (point simulation option of CLM). More information about checkout_externals can be found in the README_EXTERNALS.rst file in your project folder. 

To checkout the externals:

```{bash checkout_externals, echo=TRUE, eval=FALSE}
./manage_externals/checkout_externals
```

## Set Up for Single Point Simulations

## Set Up for Point Simulations (PTCLM)

## Setting Up and Running a Case

## Model Outputs

## Useful References

