---
title: "Reynold's Creek Experimental Watershed: Single Point Simulations"
author: "Katie Murenbeeld"
date: "07/31/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1. Introduction

Reynold's Creek Experimental Watershed (RCEW)...(a little about the location and CZO).

About the Community Earth Systems Model (CESM), Community Terrestrial Systems Model (CTSM) and Community Land Model (CLM) explain how different although terms are used almost interchangably. Model simulations. Include a link to the escomp CLM5.0 and the CLM4.5 User's Guides. 

The purpose of this technical note is to familiarize a LEAF lab user with setting up and running a single point simulation at RCEW with meterological data from one of the AmeriFlux towers on the site. In this document, I am assuming that you have a small or large allocation on the National Center for Atmospheric Research's (NCAR) supercomputer Cheyenne. At this time CLM is not ported on the high performance computer (HPC) R2 at Boise State, but this is a goal to complete soon. 

```{r clm_fig01, echo=FALSE, eval=FALSE}

```

## 2. File Structure and Cloning CTSM

### 2.1 File Structure

If you don't have a git working directory set up already on Cheyenne, do so now. It is a best practice to keep projects in their own work directory within this git working directory. 

```{bash set_dir, echo=TRUE, eval=FALSE}
cd /glade/work/$USER/ # go to your work directory on Glade
mkdir git             # create a new folder that will be the parent for future projects using CLM/CTSM
cd git                # go to the newly created git folder
```

### 2.2 Cloning CTSM

From your newly created git folder, clone the CLM/CTSM code. The following code will point to the latest CTSM release:

```{bash git_clone, echo=TRUE, eval=FALSE}
# Check out the latest release branch and create a new branch for your project

git clone -b release-clm5.0 https://github.com/ESCOMP/CTSM.git ctsm_rcew 
cd ctsm_rcew   # A new directory was created for your project
git branch   # Check to see the branches you have now created. There should be a master (or main) as well as a ctsm_rcew branch. The ctsm_rcew branch should have a star next to it meaning that is the active branch or branch you are working in.

# If no new branch, create your own.
git checkout -b ctsm_rcew
ls   # Explore the contents of the new directory
```

Notice that there is a folder within your project direcotry named manage_externals. This contains a very important script, checkout_externals, which downloads all of the components required for building and running CTSM. These include: CLM (land model component), CISM (sea-ice component), RTM (river routing component), MOSART (river routing component), CIME (contains scripts and tools for running CESM), CMEPS (Earth prediction option of CIME), FATES (dynamic vegetation option of CLM), and PTCLM (point simulation option of CLM). More information about checkout_externals can be found in the README_EXTERNALS.rst file in your project folder. 

To checkout the externals:

```{bash checkout_externals, echo=TRUE, eval=FALSE}
./manage_externals/checkout_externals
```

It is also a good idea to create separate folders for different projects in your git directory. You can either re-clone ctsm for each proect or do a recursive copy of another project.

For example, I have several projects that live in my git folder.

```{bash git_folder, echo=TRUE, eval=FALSE}
cd /glade/work/katiem/git
ls
> ctsm  ctsm_py  ctsm_sits ctsm_rcew
```


## 3. Set Up for Point Simulations (PTCLM)

The **P**oin**T** **CLM** runs through the same steps for creating single point domain and surface data files as in section 2. PTCLM is a simpler way to create these files specifically for tower sites.  The file PTCLMmkdata runs the tools to get datasets set up, and copies them to a location you can use, including the changes needed for a case to use the dataset with namelist and XML changes. There are a few steps you will need to take before you can run the PTCLMmkdata script.

First, you will need to add tower site specific data to the 

Once again load the appropriate modules in the correct order (same as below). **NOTE** Depending on if you use miniconda you may or may mot need to load all of the modules if they are already in the miniconda environment. For example, here I do not *module load python/2* because I will activate a python 2 miniconda environment later.

```{bash ptcml_modload, echo=TRUE, eval=FALSE}
module load ncarenv/1.3
module load intel/17.0.1
module load ncarcompilers/0.5.0
module load nco/4.7.9
module load netcdf/4.7.3
module load ncl/6.6.2 (H)
module load mpt/2.22
module load esmf_libs/7.1.0r (H)
module load ncview/2.1.7
```

Change directory to the PTCLM folder. From there execute the buildtools script. Buildtools will...

```{bash ptcml_folder, echo=TRUE, eval=FALSE}
CTSMROOT=/glade/work/katiem/git/ctsm_rcew
cd $CTSMROOT/tools/PTCLM
./buildtools
```

There is a known issue within the PTCLMmkdata code which will need to be addressed before you can continue on with the process. Use the text editor of your choice to open PTCLMmkdata. In line 552 you will need to remove the "export" command.

```{bash PTCLMmkdata, echo=TRUE, eval=FALSE}
- cmd = "export REGRID_PROC=1; "+mkmapdat_dir+"/mkmapdata.sh --gridfile "+scripgridfile+" --res "+clmres+" --gridtype regional -v > "+mapdir+"/mkmapdata.log";
+ cmd = mkmapdat_dir+"/mkmapdata.sh --gridfile "+scripgridfile+" --res "+clmres+" --gridtype regional -v > "+mapdir+"/mkmapdata.log";
```

Once all of the modules are loaded, the tools are built, and the PTCLMmkdata script is updated, open up a qinteractive session in order to complete the map generation and file creation process. The process can take several hours, so I suggest setting the walltime to 6 hours.

```{bash ptcml_qint, echo=TRUE, eval=FALSE}
qinteractive -X -l walltime=4:00:00
```

Load a python 2 compatible miniconda environment (optional).

```{bash ptclm_py2, echo=TRUE, eval=FALSE}
conda activate python2
```

After entering the qinteractive environment, you will need to reset environmental variables.
```{bash ptml_setenv, echo=TRUE, eval=FALSE}
CSMDATA=/glade/p/cesm/cseg/inputdata
CTSMROOT=/glade/work/katiem/git/ctsm_rcew
SITE=US-Rwe
```

Next, run the PTCLMsublist, which will... 
```{bash ptclm_sublist, echo=TRUE, eval=FALSE}
./PTCLMsublist -l $SITE -d $CSMDATA -o --verbose --account=UBOI0003 --mach=cheyenne
```

PTCLMsublist will output a command to copy and paste into the terminal. It will look something like this...
```{bash ptclm_mkdata, echo=TRUE, eval=FALSE}
./PTCLMmkdata -s US-Rwe -d /glade/p/cesm/cseg/inputdata
```

## 5. Accelerated Spin Up

Something something, working with soils, getting to a steady state. Need to do a spin up. 

Within this section, also some of the (not so basic) basics for creating and submitting a CLM "case", or simulation.

## 6. Setting Up and Running a Case

## 7. Model Outputs

## 8. Useful References

