---
title: "Reynold's Creek Experimental Watershed: Single Point Simulations"
author: "Katie Murenbeeld"
date: "6/27/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1. Introduction

Reynold's Creek Experimental Watershed (RCEW)...(a little about the location and CZO).

About the Community Earth Systems Model (CESM), Community Terrestrial Systems Model (CTSM) and Community Land Model (CLM) explain how different although terms are used almost interchangably. Model simulations. Include a link to the escomp CLM5.0 and the CLM4.5 User's Guides. 

The purpose of this technical note is to familiarize a LEAF lab user with setting up and running a single point simulation at RCEW with meterological data from one of the AmeriFlux towers on the site. In this document, I am assuming that you have a small or large allocation on the National Center for Atmospheric Research's (NCAR) supercomputer Cheyenne. At this time CLM is not ported on the high performance computer (HPC) R2 at Boise State, but this is a goal to complete soon. 

```{r clm_fig01, echo=FALSE, eval=FALSE}

```

## 2. File Structure and Cloning CTSM

### 2.1 File Structure

If you don't have a working directory set up already on Cheyenne, do so now. It is a best practice to keep projects in your work directory. 

```{bash set_dir, echo=TRUE, eval=FALSE}
cd /glade/work/$USER/ # go to your work directory on Glade
mkdir git             # create a new folder that will be the parent for future projects using CLM/CTSM
cd git                # go to the newly created git folder
```

### 2.2 Cloning CTSM

From your newly created git folder, clone the CLM/CTSM code. The following code will point to the latest CTSM release:

```{bash git_clone, echo=TRUE, eval=FALSE}
# Check out the latest release branch and create a new branch for your project

git clone -b release-clm5.0 https://github.com/ESCOMP/CTSM.git ctsm_rcew 
cd ctsm_rcew   # A new directory was created for your project
git branch   # Check to see the branches you have now created. There should be a master (or main) as well as a ctsm_rcew branch. The ctsm_rcew branch should have a star next to it meaning that is the active branch or branch you are working in.

# If no new branch, create your own.
git checkout -b ctsm_rcew
ls   # Explore the contents of the new directory
```

Notice that there is a folder within your project direcotry named manage_externals. This contains a very important script, checkout_externals, which downloads all of the components required for building and running CTSM. These include: CLM (land model component), CISM (sea-ice component), RTM (river routing component), MOSART (river routing component), CIME (contains scripts and tools for running CESM), CMEPS (Earth prediction option of CIME), FATES (dynamic vegetation option of CLM), and PTCLM (point simulation option of CLM). More information about checkout_externals can be found in the README_EXTERNALS.rst file in your project folder. 

To checkout the externals:

```{bash checkout_externals, echo=TRUE, eval=FALSE}
./manage_externals/checkout_externals
```

It is also a good idea to create separate folders for different projects in your git directory. You can either re-clone ctsm for each proect or do a recursive copy of another project.

For example, I have several projects that live in my git folder.

```{bash git_folder, echo=TRUE, eval=FALSE}
cd /glade/work/katiem/git
ls
> ctsm  ctsm_py  ctsm_sits ctsm_rcew
```

## 3. Set Up for Single Point Simulations
Note to self: Include the script used to generate the single point surface dataset. Then breakout and step through the different parts of the script. Reference = CLM5.0 UG 1.3.5 and 1.6.3.5.

```{bash spt_creation, echo=TRUE, eval=FALSE}
#!/usr/bin/env bash

#####################################################################################
## Draft script for making a single point surface and domain file for RCEW      #####
## DATE: 27 June 2020                                                           #####
## AUTHOR: Katie Murenbeeld                                                     #####
## Main code from clm5.0 User's Guide Chapter 1.6.3.5 found here:               #####
## https://escomp.github.io/ctsm-docs/versions/release-clm5.0/html/users_guide/ #####
#####################################################################################
## Before starting make sure to follow the instructions in the INSTALL file found here
## $CTSMROOT/cime/tools/mapping/gen_domain_files

# Load the correct modules
module load intel/17.0.1
module load ncl/6.6.2
#module load nco not sure if this is needed, but putting here as a reminder to test.
module load esmf_libs/7.1.0r

# Set the environmental variables
USER=katiem
CTSMROOT=/glade/work/${USER}/git/ctsm_rcew
CIMEROOT=/glade/work/${USER}/git/ctsm_rcew/cime

#GRIDNAME=10x10min_rcewID
GRIDNAME=10x10min #testing out using a different grid name. I would like to keep rcew in the name if I can.
RES=10x10min
CDATE=`date '+%y%m%d'`

# Create the SCRIP grid file for the location and create a unity mapping file for it.
# Coordinates from tower location sites:
# US-Rls = 43.1439, 243.2644
# US-Rms = 43.0645, 243.2514
# US-Rwe = 43.0653469, 243.2408977
# US-Rwf = 43.120711, 243.276914

cd $CTSMROOT/tools/mkmapdata
./mknoocnmap.pl -p 43,243 -n $GRIDNAME -v

# Set point to MAPFILE just created that will be used later

MAPFILE=$CTSMROOT/tools/mkmapdata/map_${GRIDNAME}_nooncean_to_${GRIDNAME}_nomask_aave_da_${CDATE}.nc

# Create the mapping files needed by mksurfdata_map

#cd ../.././mkmapdata
GRIDFILE=$CTSMROOT/tools/mkmapgrids/SCRIPgrid_${GRIDNAME}_nomask_c${CDATE}.nc
./mkmapdata.sh -r $RES -f $GRIDFILE -t regional

# Create the domain file

cd $CIMEROOT/tools/mapping/gen_domain_files/src
../../../configure --macros-format Makefile --mpilib mpi-serial
(. ./.env_mach_specific.sh ; gmake)
gmake
cd ../
OCNDOM=domain.ocn_noocean.nc
ATMDOM=domain.lnd.${GRIDNAME}_noocean.nc
./gen_domain -m $MAPFILE -o $OCNDOM -l $ATMDOM

# Save the location where the domain file was created

#GENDOM_PATH='pwd'

# Finally, create the surface dataset
cd $CTSMROOT/tools/mksurfdata_map/src
gmake
cd ..
./mksurfdata.pl -r usrspec -usr_gname $GRIDNAME -usr_gdate $CDATE
```

As a note, you do not need to create a script to complete this process. You can also work from the command line. In which case you would set environmental variables like so:

```{bash setenv, echo=TRUE, eval=FALSE}
USER=katiem
CTSMROOT=/glade/work/${USER}/git/ctsm_rcew
```

And then run all of the same commands as in the script above but in the command line. Now, let's break this script down. 

The first step is to load the appropriate modules (i.e. packages and libraries). The module **ncl** (NCAR command language) is required to run some of the scripts. The Earth Systems Model Framework library (**esmf_libs/7.1.0r**) is needed to run some of the mkmap scripts. In order use **esmf_lib/7.1.0r** and ncl you will need to use intel version 17.0.1 or 18.0.5. **Note!** You can easily look up which versions of modules are required by using the module spider command.

```{bash mod_spid, echo=TRUE, eval=FALSE}
module spider esmf_libs/7.1.0.
```


```{bash spt_modload, echo=TRUE, eval=FALSE}
module load ncl
module load intel/18.0.5
```


## 4. Set Up for Point Simulations (PTCLM)

The **P**oin**T** **CLM** runs through the same steps for creating single point domain and surface data files as in section 2. 

Once again load the appropriate modules in the correct order.

```{bash ptcml_modload, echo=TRUE, eval=FALSE}
module load ncarenv/1.3
module load intel/17.0.1
module load ncarcompilers/0.5.0
module load nco/4.7.9
module load netcdf/4.7.3
module load ncl/6.6.2 (H)
module load mpt/2.22
module load esmf_libs/7.1.0r (H)
module load ncview/2.1.7
```

Change directory to the PTCLM folder. From there execute the buildtools script. Buildtools will...

```{bash ptcml_folder, echo=TRUE, eval=FALSE}
CTSMROOT=/glade/work/katiem/git/ctsm_rcew
cd $CTSMROOT/tools/PTCLM
./buildtools
```

There is a known issue within the PTCLMmkdata code which will need to be addressed before you can continue on with the process. Use the text editor of your choice to open PTCLMmkdata. In line 552 you will need to remove the "export" command.

```{bash PTCLMmkdata, echo=TRUE, eval=FALSE}
- cmd = "export REGRID_PROC=1; "+mkmapdat_dir+"/mkmapdata.sh --gridfile "+scripgridfile+" --res "+clmres+" --gridtype regional -v > "+mapdir+"/mkmapdata.log";
+ cmd = mkmapdat_dir+"/mkmapdata.sh --gridfile "+scripgridfile+" --res "+clmres+" --gridtype regional -v > "+mapdir+"/mkmapdata.log";
```


## 5. Setting Up and Running a Case

## 6. Accelerated Spin Up

## 7. Model Outputs

## 8. Useful References

